### 2019-12-11 conclusions.tex
Final draft

\chapter{Conclusions} \label{chap:conclusions}
\section{Summary} \label{sec:conclusions_summary}
Several analyses have been presented of events with a monophoton signature selected out of 35.9\fbinv\ of 13\unit{TeV} \Pp\Pp\ collision
data collected by the CMS experiment at CERN in 2016. The primary expected SM contribution to such events is
the \zinvg\ process, and a measurement of the production cross section of this process at 13\unit{TeV} has been obtained at state-of-the-art
precision. In particular, the cross section measurement for events with $\pTgamma > 600\unit{GeV}$, found here to be
$1.18 \pm 0.04\mathrm{(syst)} \pm 0.36\mathrm{(stat)}$, has roughly half the percentage uncertainty of the next-most-precise
measurement in this regime, published by ATLAS in 2018~\cite{ref:CERN-EP-2018-220}. This is the first measurement of the \zinvg\ cross section
at 13\unit{TeV} reported by CMS, and the first CMS measurement of this cross section unfolded into separate bins of \pTgamma.

The results presented here do not show any $5\sigma$ deviations from SM predictions in favor of any other model.
The 95\% CL lower limit on the suppression scale $\Lambda$ of the DM-EWK EFT is 850\unit{GeV} for small \mdm\ values, the strongest
limits on this model to date. This represents an increase of 60\unit{GeV} over the strongest ATLAS limit on $\Lambda$~\cite{ref:epjc/s10052-017-4965-8},
and an increase of 250\unit{GeV} over the prior CMS results based on 12.9\fbinv\ of 2016 data~\cite{ref:JHEP10(2017)073}.
For DM simplified models, mediator masses up to 950\unit{GeV} are excluded at 95\% CL for small \mdm\ values,
also an increase of 250\unit{GeV} over the previous CMS analysis. The strongest observed monophoton limit on \mmed\ remains
1200\unit{GeV}, that of the ATLAS experiment based on their analysis of 2016 data~\cite{ref:epjc/s10052-017-4965-8}.
The 95\% CL lower limit on the ADD mass scale \mD\ is 2.85\unit{TeV} for $n = 3$ extra dimensions, rising to 2.90\unit{TeV} for $n = 6$.
This represents an increase of 410\unit{GeV} at high $n$ over the prior CMS results based on 12.9\fbinv\ of 2016 data~\cite{ref:JHEP10(2017)073},
and up to 560\unit{GeV} at low $n$. These are the strongest ADD limits in the monophoton channel to date.
The DM and ADD results presented this thesis have been published in the Journal of High Energy Physics as ``Search for new physics in final states with
a single photon and missing transverse momentum in proton-proton collisions at $\sqrt{s} = 13\unit{TeV}$''~\cite{ref:JHEP02(2019)074}.

@@

The expected limit on \mmed\ for the DM simplified models, based on prefit expected yields,
it 200\unit{GeV} higher than the observed limit. This is consistent with the excess of observed data at high \ETgamma\ compared to the prefit
expected yield (see Fig.~\ref{fig:postfitDM_SR}), and also consistent with the larger value of the expected limits on $\Lambda$ and \mD\ compared
to the observed limits (see Figs.~\ref{fig:DMEWKlimits}, \ref{fig:MDLimits}). A total of 16 observed events populate the highest \ETgamma\ bin
in both signal regions combined, and the statistical errors relating to this bin are a sizable fraction of the total event count, motivating the
conjecture that the large ratio of observed to predicted events is a statistical fluke. However, as noted in sec.~\ref{sec:results_znng_xsec}, the probability
of the \zinvg\ process alone contributing the number of observed events in the highest \ETgamma\ bin is rather small, assuming that the \zinvg\ cross
section is adequately modeled by the MATRIX NNLO predictions and that the behavior of \zinvg\ events in the CMS experiment is adequately modeled
by the MadGraph5+GEANT4 simulations. Either of these may in fact be insufficiently accurate at high \pTgamma, relative to current experimental
sensitivity. Alternatively, some unknown or inadequately modeled source of background may be making a significant contribution. The most
provocative possibility is that this represents the first subtle glimpse of BSM physics, though the estimated probability of a \zinvg\ fluctuation
is thousands of times higher than the accepted threshold to claim such a discovery. Some combination of any or all of these possibilities may be true.
Additional data would improve the statistical precision of the high-\ETgamma\ yield estimates, sharpening any potential excesses, or else
``regressing to the mean'' and leveling out to the current SM predictions.

>>>>

\chapter{Conclusions} \label{chap:conclusions}
\section{Summary} \label{sec:conclusions_summary}
Several analyses have been presented of events with a monophoton signature selected out of 35.9\fbinv\ of 13\unit{TeV} \Pp\Pp\ collision
data collected by the CMS experiment at CERN in 2016. The primary expected SM contribution to such events is
the \zinvg\ process, and a measurement of the production cross section of this process at 13\unit{TeV} has been obtained at state-of-the-art
precision. In particular, the cross section measurement for events with $\pTgamma > 600\unit{GeV}$, found here to be
$1.18 \pm 0.04\mathrm{(syst)} \pm 0.36\mathrm{(stat)}\unit{fb}$, has roughly half the percentage uncertainty of the next-most-precise
measurement in this regime, published by ATLAS in 2018~\cite{ref:CERN-EP-2018-220}. This is the first measurement of the \zinvg\ cross section
at 13\unit{TeV} reported by CMS, and the first CMS measurement of this cross section unfolded into separate bins of \pTgamma.

The results presented here do not show any $5\sigma$ deviations from SM predictions in favor of any other model.
The 95\% CL lower limit on the suppression scale $\Lambda$ of the DM--EWK EFT is 850\unit{GeV} for small \mdm\ values, the strongest
limits on this model to date. This represents an increase of 60\unit{GeV} over the strongest ATLAS limit on $\Lambda$~\cite{ref:epjc/s10052-017-4965-8},
and an increase of 250\unit{GeV} over the prior CMS results based on 12.9\fbinv\ of 2016 data~\cite{ref:JHEP10(2017)073}.
For DM simplified models, mediator masses up to 950\unit{GeV} are excluded at 95\% CL for small \mdm\ values,
also an increase of 250\unit{GeV} over the previous CMS analysis. The strongest observed monophoton limit on \mmed\ remains
1200\unit{GeV}, that of the ATLAS experiment based on their analysis of 2016 data~\cite{ref:epjc/s10052-017-4965-8}.
The 95\% CL lower limit on the ADD mass scale \mD\ is 2.85\unit{TeV} for $n = 3$ extra dimensions, rising to 2.90\unit{TeV} for $n = 6$.
This represents an increase of 410\unit{GeV} at high $n$ over the prior CMS results based on 12.9\fbinv\ of 2016 data~\cite{ref:JHEP10(2017)073},
and up to 560\unit{GeV} at low $n$. These are the strongest ADD limits in the monophoton channel to date.
The DM and ADD results presented this thesis have been published in the Journal of High Energy Physics as ``Search for new physics in final states with
a single photon and missing transverse momentum in proton-proton collisions at $\sqrt{s} = 13\unit{TeV}$''~\cite{ref:JHEP02(2019)074}.

@@

The expected limit on \mmed\ for the DM simplified models, based on prefit expected yields,
is 200\unit{GeV} higher than the observed limit. This is consistent with the excess of observed data at high \ETgamma\ compared to the prefit
expected yield (see Fig.~\ref{fig:postfitDM_SR}), and also consistent with the larger value of the expected limits on $\Lambda$ and \mD\ compared
to the observed limits (see Figs.~\ref{fig:DMEWKlimits}, \ref{fig:MDLimits}). A total of 16 observed events populate the highest \ETgamma\ bin
in both signal regions combined, and the statistical errors relating to this bin are a sizable fraction of the total event count, motivating the
conjecture that the large ratio of observed to predicted events is a statistical fluke. However, as noted in sec.~\ref{sec:results_znng_xsec}, the probability
of the \zinvg\ process alone contributing the number of observed events in the highest \ETgamma\ bin is rather small, assuming that the \zinvg\ cross
section is adequately modeled by the MATRIX NNLO predictions and that the behavior of \zinvg\ events in the CMS experiment is adequately modeled
by the MadGraph5+GEANT4 simulations. Either of these may in fact be insufficiently accurate at high \pTgamma, relative to current experimental
sensitivity. Alternatively, some unknown or inadequately modeled source of background may be making a significant contribution. The most
provocative possibility is that this represents the first subtle glimpse of BSM physics, though the estimated probability of a \zinvg\ fluctuation
is thousands of times higher than the accepted threshold to claim such a discovery. Some combination of any or all of these possibilities may be true.
Additional data would improve the statistical precision of the high-\ETgamma\ yield estimates, sharpening any potential excesses, or else
``regressing to the mean'' and leveling out to the current SM predictions.

### 2019-12-11 draft.tex
General touching up
Final draft

\settitle{A Measurement of \zinvg\ Production and a Search for New Physics in
Monophoton Events Using the CMS Detector at the LHC}
\setauthor{James Joseph Buchanan}
\setdepartment{Physics}
\doctors % or \masters
\setgraddate{2019}
\setdefensedate{16 August} % or whatever format you want

%%%% Members of the Final Oral Committee (FOC)
%%%% Give name, rank, and department
%%%% 
\setfoca{Sridhara Dasu}{Professor}{Physics} % <- Your advisor
\setfocc{Matthew Herndon}{Professor}{Physics}
\setfocd{Daniel Chung}{Professor}{Physics}
\setfoce{Jordan Ellenberg}{Professor}{Mathematics}
% \setfocf{Grover Cleveland}{Professor}{Zoology}

@@

>>>>

\settitle{A Measurement of \zinvg\ Production and a Search for New Physics in
Monophoton Events Using the CMS Detector at the LHC}
\setauthor{James Joseph Buchanan}
\setdepartment{Physics}
\doctors % or \masters
\setgraddate{2019}
\setdefensedate{16 August 2019} % or whatever format you want

%%%% Members of the Final Oral Committee (FOC)
%%%% Give name, rank, and department
%%%% 
\setfoca{Sridhara Dasu}{Professor}{Physics} % <- Your advisor
\setfocb{Matthew Herndon}{Professor}{Physics}
\setfocc{Kevin Black}{Professor}{Physics}
\setfocd{Daniel Chung}{Professor}{Physics}
\setfoce{Jordan Ellenberg}{Professor}{Mathematics}

@@

% Acknowledgements go here if you want to include them.
\section{Acknowledgements}
This document, and the work it represents, would never have come together without
the support of my colleagues and friends in the UW--Madison CMS group.
Wesley Smith, Sridhara Dasu, Matt Herndon, Tulika Bose, and Kevin Black
have nurtured a rigorous standard of excellence that the group can rightly be proud of.
Bhawna Gomber's deep expertise and untiring work have pushed these analyses
forward at every step.
Tom Perry helped to develop a previous iteration of these analyses,
and freely contributed his assistance and code to let us hit the ground running in 2016.
Devin Taylor, Tyler Ruggles, Nate Woods, Laura Dodd, Nick Smith, Kenneth Long, and
Usama Hussein were all top-notch office mates whose insights I've taken advantage of
on countless occasions.
I must finally thank my parents, Yvonne and Darryl Buchanan,
for their perpetual support of everything I do.
\clearpage

### 2019-12-11 estimation.tex
Final draft

Separate template fits are performed in four different bins of reconstructed \ETgamma.
The fits are further separated by whether the probe has a matching pixel seed. These are illustrated in Fig.~\ref{fig:efake_fits}.
After the fits, the integrals of the signal templates with and without
pixel seeds in reco bin $j$ are denoted by $N^{\Pe\Pe}[j]$ and $N^{\Pe\gamma}[j]$, respectively.
The fake ratio $R_{\Pe}[j]$ in reco bin $j$ is estimated to be $N^{\Pe\gamma}[j] / N^{\Pe\Pe}[j]$.

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=0.40\linewidth]{Figures/efake/EE175190.png}
    \includegraphics[width=0.40\linewidth]{Figures/efake/EP175190.png}
    \includegraphics[width=0.40\linewidth]{Figures/efake/EE190220.png}
    \includegraphics[width=0.40\linewidth]{Figures/efake/EP190220.png}
    \includegraphics[width=0.40\linewidth]{Figures/efake/EE220270.png}
    \includegraphics[width=0.40\linewidth]{Figures/efake/EP220270.png}
    \includegraphics[width=0.40\linewidth]{Figures/efake/EE270.png}
    \includegraphics[width=0.40\linewidth]{Figures/efake/EP270.png}
    \caption{
      Fits to the mass distributions for $\Pe\Pe$ (left) and $\Pe\gamma$ (right) selections, in bins of probe \pT: $175 < \pT < 190\unit{GeV}$ , $190 < \pT < 220\unit{GeV}$ , $220 < \pT > 270\unit{GeV}$  and $ \pT > 270\unit{GeV}$.
      The blue dashed curve is the background template, and the red solid curve is the sum of the signal and background templates.
    }
    \label{fig:efake_fits}
  \end{center}
\end{figure}

@@

The fake ratio is esimated by counting the number of numerator and denominator events in a control region defined by $\MET < 30\unit{GeV}$. Both the observed
numerator and denominator yields are expected to receive ``contamination'' from real hard-process photon events. The expected contribution of real hard-process
photons to the observed numerator yield is estimated (and subtracted out) by fitting real and fake photon templates to the photon \sieie\ distribution of numerator
events. The real photon template is estimated using \gjets\ MC. The fake photon template is derived from observed event yields in a \textit{sideband} region in which
the leading photon has worst charged hadron isolation lying between 8 and 14\unit{GeV}, but otherwise passes numerator cuts.
This sideband was chosen for the relative absence of real photon contamination. Nevertheless, some residual real photon events are still expected to contribute, and
the fake template is finally ``cleaned'' by subtracting the expected \sieie\ distribution of real photon events estimated in \gjets\ MC.

@@

\begin{table}[tbp]
  \begin{center}
    \label{tab:response_matrix_values}
    \begin{tabular}{c c c c c c}
      \hline
       0.297401 & 0.0428149 & 0.000595967 & 5.87982e-05 & 0 & 0 \\
       0.0399934 & 0.38611 & 0.049973 & 0.000505446 & 0.000133614 & 0 \\
       0.000202488 & 0.0214437 & 0.37027 & 0.0326231 & 0.000264926 & 0.000183209 \\
       0 & 0.000196144 & 0.0229108 & 0.367465 & 0.0249473 & 0 \\
       0 & 0 & 1.50756e-05 & 0.0100609 & 0.343767 & 0.022748 \\
       0 & 0 & 0 & 0 & 0.00326465 & 0.316482 \\
      \hline
    \end{tabular}
    \caption{
      The values of the response matrix $K$, estimated in \zinvg\ MadGraph MC with higher order corrections applied.
      With rows ordered from top to bottom and columns ordered from left to right, the value in row $i$, column $j$
      is equal to the conditional probability of a \zinvg\ event to pass all signal region cuts and land in reconstructed
      \ETgamma\ bin $j$, if its true \pTgamma\ value is in bin $i$.
    }
  \end{center}
\end{table}

@@

The complete likelihood function that is maximized in the SM cross section fit is
\begin{align}
\begin{split}
  \mathcal{L} & = \left\{ \mathcal{L_{\text{Signal}}} \mathcal{L_{\text{Control}}} \right\} \times \mathcal{L_{\text{Nuisances}}} \\
  & = \prod_{j} \left\{
    \prod_{S=H,V} \mathcal{P}\left( N_{S, j} \left| \nS[,j] (\vec{\theta} \right.) \right) \prod_{\ell=\Pe,\Pmu} \mathcal{P}\left( N_{\ell\Pgamma, j} \left| \nl[,j] (\vec{\theta}) \right. \right) 
    \right\}  \times \prod_{t} \mathcal{N}(\theta_t) \\
  & = \prod_{j} \left\{
  \begin{array}{cc}
    \prod\limits_{S=H,V} \mathcal{P}\left( N_{S, j} \left| C_{S} \left(\Sigma_{i}r_{i}\nZg[j,i](\vec{\theta}) + \nWg[j] + b_{j}(\vec{\theta})\right) + h \nhalo[,j](\vec{\theta}) \right. \right) \\[1.3pt]
    \times \prod\limits_{\ell=\Pe,\Pmu} \mathcal{P}\left( N_{\ell\Pgamma, j} \left| \frac{\nWg[j]}{\RWl[,j](\vec{\theta})} + b_{\ell\Pgamma, j}(\vec{\theta}) \right. \right) \\[1.3pt]
  \end{array} \right\}
  \times \prod_{t} \mathcal{N}(\theta_t)
\end{split}
\end{align}
following the notation introduced above, and where $\mathcal{P}(N\vert\lambda)$ is the Poisson probability of $N$ for mean $\lambda$,
$\mathcal{N}$ denotes the unit normal distribution centered at zero, and $N_{X,i}$ is the observed number of events in reco bin i of region X.

@@

The total reconstructed event yield \nS\ in each signal region is estimated according to
\begin{equation}
  \nS[,j] = C_{S} (\mu s_{j} + (1 + \fZW[,j])\nZg[j,all] + b_{j}) + h\nhalo[,j]
\end{equation}
where $\nZg[j,all]$ refers to the total contribution from all \zinvg\ events to reco \ETgamma\ bin $j$, $s_{j}$ is the contribution from the BSM
signal to reco bin $j$ (of both signal regions combined), and $\mu$ is a constant that represents the overall BSM signal strength.
Instead of being estimated in MC, \nZg[j,all] is a freely floating fit parameter (one for each reco \ETgamma\ bin $j$).
The total reconstructed event yield \nl\ in each single-$\ell$ control region ($\ell \in \{e,\mu\}$) is estimated according to
\begin{equation}
  \nl[,j] = \frac{\fZW[,j]\nZg[j,all]}{\RWl[,j]} + b_{\ell\Pgamma,j}
\end{equation}
The total reconstructed event yield \nll\ in each double-$\ell$ control region is similarly estimated according to
\begin{equation}
  \nll[,j] = \frac{\nZg[j,all]}{\RZll[,j]} + b_{\ell\ell\Pgamma,j}
\end{equation}
where $b_{\ell\ell\Pgamma,j}$ is the estimated contribution from processes other than \zllg\ in the double-$\ell$ control region.
The complete likelihood function is
\begin{align}
\begin{split}
  \mathcal{L} & = \left\{ \mathcal{L_{\text{Signal}}} \mathcal{L_{\text{Control}}} \right\} \times \mathcal{L_{\text{Nuisances}}} \\
  & = \prod_{j} \left\{
    \prod_{S=H,V} \mathcal{P}\left( N_{S, j} \left| \nS[,j] (\vec{\theta} \right.) \right)
     \prod_{\ell=\Pe,\Pmu} \mathcal{P}\left( N_{\ell\Pgamma, j} \left| \nl[,j] (\vec{\theta}) \right. \right)
      \prod_{\ell=\Pe,\Pmu} \mathcal{P}\left( N_{\ell\ell\Pgamma, j} \left| \nll[,j] (\vec{\theta}) \right. \right) 
    \right\}  \times \prod_{t} \mathcal{N}(\theta_t) \\
  & = \prod_{j} \left\{
  \begin{array}{cc}
    \prod\limits_{S=H,V} \mathcal{P}\left( N_{S, j} \left| C_{S} \left(\mu s_{j}(\vec{\theta}) + (1 + \fZW[,j])\nZg[j,all](\vec{\theta}) + b_{j}(\vec{\theta})\right) + h \nhalo[,j](\vec{\theta}) \right. \right) \\[1.3pt]
    \times \prod\limits_{\ell=\Pe,\Pmu} \mathcal{P}\left( N_{\ell\Pgamma, j} \left| \frac{\fZW[,j]\nZg[j,all]}{\RWl[,j](\vec{\theta})} + b_{\ell\Pgamma, j}(\vec{\theta}) \right. \right) \\[1.3pt]
    \times \prod\limits_{\ell=\Pe,\Pmu} \mathcal{P}\left( N_{\ell\ell\Pgamma, j} \left| \frac{\nZg[j,all]}{\RZll[,j](\vec{\theta})} + b_{\ell\ell\Pgamma, j}(\vec{\theta}) \right. \right) \\[1.3pt]
  \end{array} \right\}
  \times \prod_{t} \mathcal{N}(\theta_t)
\end{split}
\end{align}

>>>>

Separate template fits are performed in four different bins of reconstructed \ETgamma.
The fits are further separated by whether the probe has a matching pixel seed. These are illustrated in Figs.~\ref{fig:efake_fits_a}, \ref{fig:efake_fits_b}.
After the fits, the integrals of the signal templates with and without
pixel seeds in reco bin $j$ are denoted by $N^{\Pe\Pe}[j]$ and $N^{\Pe\gamma}[j]$, respectively.
The fake ratio $R_{\Pe}[j]$ in reco bin $j$ is estimated to be $N^{\Pe\gamma}[j] / N^{\Pe\Pe}[j]$.

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=0.40\linewidth]{Figures/efake/EE175190.png}
    \includegraphics[width=0.40\linewidth]{Figures/efake/EP175190.png}
    \includegraphics[width=0.40\linewidth]{Figures/efake/EE190220.png}
    \includegraphics[width=0.40\linewidth]{Figures/efake/EP190220.png}
    \caption{
      Fits to the mass distributions for $\Pe\Pe$ (left) and $\Pe\gamma$ (right) selections, for probe \pT\ between 175 and 190\unit{GeV} (upper) and
      between 190 and 220\unit{GeV} (lower). The blue dashed curve is the background template, and the red solid curve is the sum of the signal and background templates.
    }
    \label{fig:efake_fits_a}
  \end{center}
\end{figure}

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=0.40\linewidth]{Figures/efake/EE220270.png}
    \includegraphics[width=0.40\linewidth]{Figures/efake/EP220270.png}
    \includegraphics[width=0.40\linewidth]{Figures/efake/EE270.png}
    \includegraphics[width=0.40\linewidth]{Figures/efake/EP270.png}
    \caption{
      Fits to the mass distributions for $\Pe\Pe$ (left) and $\Pe\gamma$ (right) selections, for probe \pT\ between 220 and 270\unit{GeV} (upper) and
      above 270\unit{GeV} (lower). The blue dashed curve is the background template, and the red solid curve is the sum of the signal and background templates.
    }
    \label{fig:efake_fits_b}
  \end{center}
\end{figure}

@@

The fake ratio is esimated by counting the number of numerator and denominator events in a control region defined by $\MET < 30\unit{GeV}$. Both the observed
numerator and denominator yields are expected to receive ``contamination'' from real hard-process photon events. The expected contribution of real hard-process
photons to the observed numerator yield is estimated (and subtracted out) by fitting real and fake photon templates to the photon \sieie\ distribution of numerator
events. The real photon template is estimated using \gjets\ MC. The fake photon template is derived from observed event yields in a \textit{sideband} region in which
the leading photon has worst charged hadron isolation lying between 8 and 14\unit{GeV}, but otherwise passes numerator cuts.
This sideband was chosen for the relative absence of real photon contamination. Nevertheless, some residual real photon events are still expected to contribute, and
the fake template is finally ``cleaned'' by subtracting the expected \sieie\ distribution of real photon events estimated in \gjets\ MC. The fits of real and fake
templates to observed data in various bins of \pTgamma\ are shown in Figs.~\ref{fig:qcd_fits_a}, \ref{fig:qcd_fits_b}, \ref{fig:qcd_fits_c}.

@@

\begin{table}[tbp]
  \begin{center}
    \label{tab:response_matrix_values}
    \begin{tabular}{c c c c c c}
      \hline
       0.350462 & 0.0278914 & 0.00032498 & 7.92826e-05 & 0 & 0 \\
       0.020877 & 0.42256 & 0.0337793 & 0.000353428 & 0.000181572 & 0 \\
       0 & 0.0111244 & 0.404499 & 0.0245065 & 0.000262814 & 0.0001816 \\
       0 & 0 & 0.0130199 & 0.386613 & 0.019395 & 0 \\
       0 & 0 & 0 & 0.00569542 & 0.352939 & 0.0204245 \\
       0 & 0 & 0 & 0 & 0.00235023 & 0.31921 \\
      \hline
    \end{tabular}
    \caption{
      The values of the response matrix $K$, estimated in \zinvg\ MadGraph MC with higher order corrections applied.
      With rows ordered from top to bottom and columns ordered from left to right, the value in row $i$, column $j$
      is equal to the conditional probability of a \zinvg\ event to pass all signal region cuts and land in reconstructed
      \ETgamma\ bin $j$, if its true \pTgamma\ value is in bin $i$.
    }
  \end{center}
\end{table}

@@

The complete likelihood function that is maximized in the SM cross section fit is
\begin{equation}
\mathcal{L} = \mathcal{L_{\text{Signal}}} \times \mathcal{L_{\text{Control}}} \times \mathcal{L_{\text{Nuisances}}}
\end{equation}
where
\begin{equation}
\mathcal{L_{\text{Nuisances}}} = \prod_{t} \mathcal{N}(\theta_t)
\end{equation}
and
\begin{align}
\begin{split}
  \mathcal{L_{\text{Signal}}} \times \mathcal{L_{\text{Control}}} & = \prod_{j} \left\{
    \prod_{S=H,V} \mathcal{P}\left( N_{S, j} \left| \nS[,j] (\vec{\theta} \right.) \right) \prod_{\ell=\Pe,\Pmu} \mathcal{P}\left( N_{\ell\Pgamma, j} \left| \nl[,j] (\vec{\theta}) \right. \right) 
    \right\} \\
  & = \prod_{j} \left\{
  \begin{array}{cc}
    \prod\limits_{S=H,V} \mathcal{P}\bigg( N_{S, j} \left| C_{S} \left(\Sigma_{i}r_{i}\nZg[j,i](\vec{\theta}) + \nWg[j] + b_{j}(\vec{\theta})\right) \\[1.3pt]
    {} + h \nhalo[,j](\vec{\theta}) \right. \bigg) \\[1.3pt]
    \times \prod\limits_{\ell=\Pe,\Pmu} \mathcal{P}\left( N_{\ell\Pgamma, j} \left| \frac{\nWg[j]}{\RWl[,j](\vec{\theta})} + b_{\ell\Pgamma, j}(\vec{\theta}) \right. \right) \\[1.3pt]
  \end{array} \right\}
\end{split}
\end{align}
following the notation introduced above, and where $\mathcal{P}(N\vert\lambda)$ is the Poisson probability of $N$ for mean $\lambda$,
$\mathcal{N}$ denotes the unit normal distribution centered at zero, and $N_{X,i}$ is the observed number of events in reco bin i of region X.

@@

The total reconstructed event yield \nS\ in each signal region is estimated according to
\begin{equation}
  \nS[,j] = C_{S} (\mu s_{j} + (1 + \fZW[,j])\nZg[j,all] + b_{j}) + h\nhalo[,j]
\end{equation}
where $\nZg[j,all]$ refers to the total contribution from all \zinvg\ events to reco \ETgamma\ bin $j$, $s_{j}$ is the contribution from the BSM
signal to reco bin $j$ (of both signal regions combined), and $\mu$ is a constant that represents the overall BSM signal strength.
Instead of being estimated in MC, \nZg[j,all] is a freely floating fit parameter (one for each reco \ETgamma\ bin $j$).
The total reconstructed event yield \nl\ in each single-$\ell$ control region ($\ell \in \{e,\mu\}$) is estimated according to
\begin{equation}
  \nl[,j] = \frac{\fZW[,j]\nZg[j,all]}{\RWl[,j]} + b_{\ell\Pgamma,j}
\end{equation}
The total reconstructed event yield \nll\ in each double-$\ell$ control region is similarly estimated according to
\begin{equation}
  \nll[,j] = \frac{\nZg[j,all]}{\RZll[,j]} + b_{\ell\ell\Pgamma,j}
\end{equation}
where $b_{\ell\ell\Pgamma,j}$ is the estimated contribution from processes other than \zllg\ in the double-$\ell$ control region.
The complete likelihood function is
\begin{equation}
\mathcal{L} = \mathcal{L_{\text{Signal}}} \times \mathcal{L_{\text{Control}}} \times \mathcal{L_{\text{Nuisances}}}
\end{equation}
where
\begin{align}
\begin{split}
  \mathcal{L_{\text{Signal}}} \times \mathcal{L_{\text{Control}}} & = \prod_{j} \left\{
  \begin{array}{cc}
    \prod_{S=H,V} \mathcal{P}\left( N_{S, j} \left| \nS[,j] (\vec{\theta} \right.) \right) \\[1.3pt]
     \times \prod_{\ell=\Pe,\Pmu} \mathcal{P}\left( N_{\ell\Pgamma, j} \left| \nl[,j] (\vec{\theta}) \right. \right) \\[1.3pt]
      \times \prod_{\ell=\Pe,\Pmu} \mathcal{P}\left( N_{\ell\ell\Pgamma, j} \left| \nll[,j] (\vec{\theta}) \right. \right)  \\[1.3pt]
  \end{array} \right\} \\
  & = \prod_{j} \left\{
  \begin{array}{cc}
    \prod\limits_{S=H,V} \mathcal{P}\bigg( N_{S, j} \left| C_{S} \left(\mu s_{j}(\vec{\theta}) + (1 + \fZW[,j])\nZg[j,all](\vec{\theta}) + b_{j}(\vec{\theta})\right) \\[1.3pt]
     {} + h \nhalo[,j](\vec{\theta}) \right. \bigg) \\[1.3pt]
    \times \prod\limits_{\ell=\Pe,\Pmu} \mathcal{P}\left( N_{\ell\Pgamma, j} \left| \frac{\fZW[,j]\nZg[j,all]}{\RWl[,j](\vec{\theta})} + b_{\ell\Pgamma, j}(\vec{\theta}) \right. \right) \\[1.3pt]
    \times \prod\limits_{\ell=\Pe,\Pmu} \mathcal{P}\left( N_{\ell\ell\Pgamma, j} \left| \frac{\nZg[j,all]}{\RZll[,j](\vec{\theta})} + b_{\ell\ell\Pgamma, j}(\vec{\theta}) \right. \right) \\[1.3pt]
  \end{array} \right\}
\end{split}
\end{align}

### 2019-12-11 experiment.tex
Final draft

The eight arc sections are each filled with a series of dipole bending magnets, 1232 in all. These produce a uniform magnetic field in a fixed direction when a current is run through them.
A magnetic field causes a charged particle's trajectory to bend, and the dipoles are arranged so as to keep the proton beam trajectories aligned with the LHC beam pipes.
Since a fixed magnetic field causes protons moving in opposite directions to bend in opposite directions, the two counter-rotating beams must be deflected by oppositely-aligned fields.
A special two-in-one magnet design was developed for the LHC that produces two side-by-side dipole fields, coupled to the same current, oriented in opposite directions. The use of
superconducting NbTi cables allows more than 12\unit{kA} to flow when the wire is cooled to 1.9\unit{K}, producing a dipole field of up to 9\unit{T}. Sophisticated cryogenic and power
transmission systems are required to operate the magnets at this level, and the two-in-one design prevents each pipe from needing its own separate system along more than 18\unit{km} of
total dipole magnet length, substantially reducing the space and cost requirements of the LHC.

@@

Differences in a related quantity called the \textit{rapidity}, defined as
\begin{equation}
\mathrm{rapidity} = \frac{1}{2}\ln{\frac{E+p_\mathrm{L}}{E-p_{L}}}
\label{eq:rapidity}
\end{equation}
are Lorentz invariant. Particle production in LHC collisions is roughly constant as a function of rapidity~\cite{ref:lhcrun1harvest_ch3}, and because
differences in rapidity do not depend on the longitudinal speed at which one is moving, that statement is independent
of whether the collisions are observed in their own rest frame or in that of the CMS. At very high
energies, rapidity becomes approximatey equal to the \textit{pseudorapidity} $\eta$ defined by
\begin{equation}
\eta = -\ln{\tan{(\theta/2)}}
\label{eq:pseudorapidity}
\end{equation}

@@

The CMS has approximately hermetic coverage of particles emerging from the IP with $|\eta|$
up to 5.0, giving a total polar coverage of 178.46 degrees along with complete 360-degree azimuthal coverage. Particles emerging
within this zone of coverage are detected with high probability by at least one subsystem, although the measurement quality varies
as a function of $|\eta|$. Since particles are emitted with a roughly constant concentration as a function of $\eta$, and because
cones of constant $\eta$ become more densely packed at higher values of $|\eta|$, it follows that the intensity of particle radiation
impinging on detector elements is stronger at higher $|\eta|$. Tradeoffs must be made between the measurement quality of an apparatus
and its radiation hardness, and therefore equipment installed at higher $|\eta|$ is typically less precise.

>>>>

The eight arc sections are each filled with a series of dipole bending magnets, 1232 in all. When a current is run through them, these produce a uniform magnetic
field inside their segments of the beam pipes, pointing perpendicular to the beam direction.
A magnetic field causes a charged particle's trajectory to bend, and the dipoles are arranged so as to keep the proton beam trajectories aligned with the beam pipes
throughout the arcs.
Since a fixed magnetic field causes protons moving in opposite directions to bend in opposite directions, the two counter-rotating beams must be deflected by oppositely-aligned fields.
A special two-in-one magnet design was developed for the LHC that produces two side-by-side dipole fields, coupled to the same current, oriented in opposite directions. The use of
superconducting NbTi cables allows more than 12\unit{kA} to flow when the wire is cooled to 1.9\unit{K}, producing a dipole field of up to 9\unit{T}. Sophisticated cryogenic and power
transmission systems are required to operate the magnets at this level, and the two-in-one design prevents each pipe from needing its own separate system along more than 18\unit{km} of
total dipole magnet length, substantially reducing the space and cost requirements of the LHC.

@@

A particle's \textit{rapidity\ y}, defined as
\begin{equation}
y = \frac{1}{2}\ln{\frac{E+p_\mathrm{L}}{E-p_{L}}},
\label{eq:rapidity}
\end{equation}
transforms as
\begin{equation}
y \rightarrow y - \tanh^{-1}\beta
\label{eq:raptransform}
\end{equation}
under a boost $\beta$ in the longitudinal velocity of one's reference frame, and so differences in $y$ are Lorentz invariant under such a
shift \cite[Ch.47]{ref:PDG}. Particle production in LHC collisions is roughly constant as a function of
rapidity~\cite{ref:lhcrun1harvest_ch3}, and because differences in rapidity do not depend on the longitudinal speed at which one is moving,
that statement is independent of whether the collisions are observed in their own rest frame or in that of the CMS. At very high
energies ($E \approx p$), rapidity becomes approximatey equal to the \textit{pseudorapidity} $\eta$ defined by
\begin{equation}
\eta = -\ln{\tan{(\theta/2)}}
\label{eq:pseudorapidity}
\end{equation}

@@

The CMS has approximately hermetic coverage of particles emerging from the IP with $|\eta|$
up to 5.0, giving a total polar coverage of 178.46 degrees along with complete 360-degree azimuthal coverage. Particles emerging
within this zone of coverage are detected with high probability by at least one subsystem, although the measurement quality varies
as a function of $|\eta|$. Since particles are emitted with a roughly constant concentration as a function of $\eta$, and because
constant differences in $\eta$ correspond to smaller $\theta$ separations at higher values of $|\eta|$, it follows that the intensity of particle radiation
impinging on detector elements is stronger at higher $|\eta|$. Tradeoffs must be made between the measurement quality of an apparatus
and its radiation hardness, and therefore equipment installed at higher $|\eta|$ is typically less precise.
